---
runtime:  shiny
title: "Embedding Shiny into RMarkdown"
author: "Mike Silva"
html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
---
```{r, echo=FALSE}
# Preview V1 at https://mikesilva.shinyapps.io/top-rockers/
```

## Top Rockers


Walt Hickey of FiveThirtyEight collected data from 25 classic rock stations operating in 30 of the largest metros for a week in June of 2014.  The data is made available to the public at https://github.com/fivethirtyeight/data/tree/master/classic-rock.

```{r setup, cache=TRUE, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(shiny)

df <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/classic-rock/classic-rock-raw-data.csv")
```

This recipie create an interative document using *shiny* and the *dpylr* and *ggplot* packages from the tidyverse.  It will allow the user to visualize which artist/group get the most airtime.  They can decide how many artists to show and filter the list by the radio station.

```{r shiny, echo=FALSE}
theme_set(theme_classic())

# Function to create the viz title
get_title <-function(CALLSIGN){
  switch(CALLSIGN,
         KCFX={return("KCFX - Kansas City, MO")},
         KGB ={return("KGB  - San Diego, CA")},
         KGLK={return("KGLK - Houston, TX")},
         KGON={return("KGLK - Portland, OR")},
         KLOS={return("KLOS - Los Angeles, CA")},
         KRFX={return("KRFX - Denver, CO")},
         KSEG={return("KSEG - Sacramento, CA")},
         KSHE={return("KSHE - St. Louis, MO")},
         KSLX={return("KSLX - Phoenix, AZ")},
         KUFX={return("KUFX - San Jose, CA")},
         KZEP={return("KZEP - San Antonio, TX")},
         KZOK={return("KZOK - Seattle, WA")},
         KZPS={return("KZPS - Dallas/Fort Worth, TX")},
         KZPS={return("KZPS - Dallas/Fort Worth, TX")},
         WAXQ={return("WAXQ - New York City, NY")},
         WBGG={return("WBGG - Ft. Lauderdale, FL")},
         WBIG={return("WBIG - Washington, DC")},
         WCSX={return("WCSX - Detroit, MI")},
         WDVE={return("WDVE - Pittsburgh, PA")},
         WLUP={return("WLUP - Chicago, IL")},
         WMGK={return("WMGK - Philadelphia, PA")},
         WNCX={return("WNCX - Cleveland, OH")},
         WOFX={return("WOFX - Cincinnati, OH")},
         WRFX={return("WRFX - Charlotte, NC")},
         WXGL={return("WXGL - Tampa Bay, FL")},
         WZLX={return("WZLX - Boston, MA")},
         {return("All Stations in the Data Set")}
  )
}

# Get the list of Radio Stations
call_signs <- df %>%
  select(CALLSIGN) %>%
  arrange(CALLSIGN) %>%
  unique() 
call_signs <- c("All Stations", as.vector(call_signs$CALLSIGN))

# Give the user a slider to pick how many artists to see.
sliderInput("n", "Number of Artist:", 10, min = 5, max = 25)

# Show the dropdown menu
selectInput("c", label = "Radio Station", choices = call_signs)

renderPlot({
  # Wrangle the data
  viz_data <- df %>%
    rowwise() %>%
    mutate(keep = ifelse(input$c=="All Stations", 1, ifelse(CALLSIGN == input$c, 1, 0))) %>%
    ungroup() %>%
    filter(keep == 1) %>%
    group_by(ARTIST.CLEAN) %>%
    summarise(Played = n()) %>%
    top_n(input$n, Played) %>%
    rename(Artist = ARTIST.CLEAN)
  
  g <- ggplot(viz_data, aes(reorder(Artist, Played), Played)) +
    geom_bar(stat="identity", width = 0.5, fill="tomato2") + 
    coord_flip() + 
    labs(caption="Source: FiveThirtyEight", 
         x = "Artist/Group Name",  
         y = "Times Songs Played on the Radio", 
         title = get_title(input$c))
  print(g)
})
```

