---
title: "Airline Safety Data"
author: "Adam Douglas"
date: "11/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Airline Data

These data are courtesy of [fivethirtyeight](https://www.fivethirtyeight.com) and relates to their article titled, [*Should Travelers Avoid Flying Airlines That Have Had Crashes in the Past?*](http://fivethirtyeight.com/features/should-travelers-avoid-flying-airlines-that-have-had-crashes-in-the-past/).

### Loading the Data
The data are stored on [fivethirtyeight's github](https://github.com/fivethirtyeight/data/tree/master/airline-safety), so we will load it directly from that site:

```{r load data, cache=TRUE}
# Load from the Github URL
airline <- read_csv(url("https://raw.githubusercontent.com/fivethirtyeight/data/master/airline-safety/airline-safety.csv"),col_types="cdiiiiii")

# Look at the data
glimpse(airline)
```

We have 8 variables in the raw data. Looking at them, however, we see that the data is not in a tidy format. Specifically, the incidents, accidents, and fatalities columns are variables, and the year ranges are observations.

We can correct this quite easily using the `tidyr` package included as part of the `tidyverse`. Specifically, we want to use the `gather` function which, as its name implies, gathers columns into rows.

```{r gather}
air <- airline %>% gather(key = "measure", value = "val", -airline, -avail_seat_km_per_week)

glimpse(air)
```

The above statement takes the columns we select (here, we selected all except `airline` and `avail_seat_km_per_week`) and transforms them so that the name of that column is populated in a new "key" column named `measure` and it's value is put into a new column named `val`.

However we're not exactly where we want to be, as incidents, fatal accidents, and fatalities are all variables and should be in their own columns.

```{r add_years}
# First split years away from the variables
air$year <- str_sub(air$measure,-5,-1)
air$measure <- str_extract(air$measure,"[a-z|_]+(?![0-9])")

# Now "spread" those variables into their own columns
airTidy <- spread(air, measure, val)

glimpse(airTidy)
```

Now it appears our dataset is tidy. Each row is an observation (airline and year range) and each column is a variable (seats per km, fatal accidents, fatalities, and incidents).
